\chapter{Appendix}
\section{Default Settings}
\label{sec:app_default_settings}
Table~\ref{tab:default_settings} lists the values for a selection of the most
important parameters in the default settings. Below, some of the ones that
aren't quite self-explanatory are explained more closely. All possible
parameters are available in the code repository in the commented default
settings file \texttt{simulation/default.settings}, relative to the repository
root (TODO check).

- TODO include decay or not?
\begin{longtable}{|l|l|}
\hline
\rowcolor{slightgray} \T Key & Value \B\\
\hline
\cellcolor{slightgray}\T \texttt{rng\_seed} & 5803379951609632196 \B\\
\hline
\cellcolor{slightgray}\T \texttt{peer\_batches} & \texttt{[(0, 64, \{\})]} \B\\
\hline
\cellcolor{slightgray}\T \texttt{num\_random\_introductions} & 8 \B\\
\hline
\cellcolor{slightgray}\T \texttt{even\_sync\_groups} & \texttt{False} \B\\
\hline
\cellcolor{slightgray}\T \texttt{initial\_reputation} & 0 \B\\
\hline
\cellcolor{slightgray}\T \texttt{force\_one\_group} & \texttt{False} \B\\
\hline
\cellcolor{slightgray}\T \texttt{ensure\_non\_empty\_sync\_groups} & \texttt{True} \B\\
\hline
\cellcolor{slightgray}\T \texttt{id\_length} & 16 \B\\
\hline
\cellcolor{slightgray}\T \texttt{prefix\_length} & 4 \B\\
\hline
\cellcolor{slightgray}\T \texttt{ignore\_non\_existent\_subprefixes} & \texttt{False} \B\\
\hline
\cellcolor{slightgray}\T \texttt{transmission\_delay} & 0.1 \B\\
\hline
\cellcolor{slightgray}\T \texttt{successful\_query\_reward} & 1 \B\\
\hline
\cellcolor{slightgray}\T \texttt{failed\_query\_penalty} & -1 \B\\
\hline
\cellcolor{slightgray}\T \texttt{timeout\_query\_penalty} & -2 \B\\
\hline
\cellcolor{slightgray}\T \texttt{no\_penalty\_reputation} & 10 \B\\
\hline
\cellcolor{slightgray}\T \texttt{decay\_timestep} & \texttt{inf} \B\\% (TODO necessary)
\hline
\cellcolor{slightgray}\T \texttt{decay\_per\_time\_unit} & 0 \B\\% (TODO necessary)
\hline
\cellcolor{slightgray}\T \texttt{reward\_attenuation $\rightarrow$ type} & \texttt{exponential} \B\\
\hline
\cellcolor{slightgray}\T \texttt{reward\_attenuation $\rightarrow$ exponent} & 0.35 \B\\
\hline
\cellcolor{slightgray}\T \texttt{reward\_attenuation $\rightarrow$ coefficient} & 1 \B\\
\hline
\cellcolor{slightgray}\T \texttt{reward\_attenuation $\rightarrow$ lower\_bound} & 10 \B\\
\hline
\cellcolor{slightgray}\T \texttt{reward\_attenuation $\rightarrow$ upper\_bound} & 15 \B\\
\hline
\cellcolor{slightgray}\T \texttt{reputation\_buffer} & 4 \B\\
\hline
\cellcolor{slightgray}\T \texttt{query\_peer\_selection} & \texttt{overlap\_rep\_sorted} \B\\
\hline
\cellcolor{slightgray}\T \texttt{query\_sync\_for\_subprefixes} & \texttt{False} \B\\
\hline
\cellcolor{slightgray}\T \texttt{uncovered\_subprefix\_interval} & 10 \B\\
\hline
\cellcolor{slightgray}\T \texttt{min\_desired\_query\_peers} & 2 \B\\
\hline
\cellcolor{slightgray}\T \texttt{max\_desired\_group\_size} & 16 \B\\
\hline
\cellcolor{slightgray}\T \texttt{query\_timeout} & 2 \B\\
\hline
\cellcolor{slightgray}\T \texttt{expect\_penalties} & \texttt{True} \B\\
\hline
\caption{Default values of a selection of settings.}
\label{tab:default_settings}
\end{longtable}

\begin{description}
\item[\texttt{peer\_batches}] A list of tuples \texttt{(time, number\_of\_peers,
settings)} that allows to add peers at different times and with individual
settings.
\item[\texttt{num\_random\_introductions}] Upon initialization, each peer is
introduced to a random sample of peers. This is the size of that sample.
\item[\texttt{even\_sync\_groups}] If \texttt{True}, initialization will take
care to make all sync groups the same size. This can be helpful in ruling out
unevenly sized sync groups as the cause of some observed effect.
\item[\texttt{force\_one\_group}] If \texttt{True}, all peers will be forced
into one query group. Can be helpful e.g. to rule out the recursive query
problem as the cause of some effect.
\item[\texttt{ensure\_non\_empty\_sync\_groups}] If \texttt{True},
initialization takes care that each sync group contains at least one peer.
\item[\texttt{prefix\_length}] Length of the routing prefix, which is just
called prefix in the simulation.
\item[\texttt{ignore\_non\_existent\_subprefixes}] If \texttt{True}, peers make
use of global information they are given in the beginning in order to not query
for subprefixes for which there doesn't exist a peer (i.e. the sync group is
empty) when they are completing their subprefix coverage. This is useful in
preventing ongoing querying for IDs that will never lead to a successful
response.
\item[\texttt{no\_penalty\_reputation}] The penalty threshold reputation.
\item[\texttt{decay\_timestep}] The interval in which the reputation decay
process is executed, which subtracts reputation from all peers. (TODO
necessary?)
\item[\texttt{decay\_per\_time\_unit}] The amount by which reputation decays
every second. (TODO necessary?)
\item[\texttt{reputation\_buffer}] The amount of reputation above the penalty
threshold peers need to have in order to consider themselves reputation
saturated.
\item[\texttt{query\_sync\_for\_subprefixes}] If \texttt{True}, peers will also
send queries to their sync peers (even if they are not query peers) when they
are completing their subprefix coverage.
\item[\texttt{uncovered\_subprefix\_interval}] Peers repeatedly check whether
they have complete subprefix coverage and send queries in order to complete it
if they don't. This is the time interval between checking.
\item[\texttt{expect\_penalties}] If \texttt{True}, peers expect to be penalized
when they perform a defecting action.
\end{description}
