\chapter{Introduction}
- TODO update submission date?

\section{Privacy in Name-Lookup}
The domain name system (DNS) is an service essential for the Internet, its
primary purpose being the mapping of domain names to IP addresses. In general, a
name mapping service that allows looking up the network address associated with
a name belonging to some entity (e.g. a server, or a user identity) is
indispensable for any large network. Using the network addresses themselves is
impractical because they are not easy to remember, and they may not be
permanently assigned.

The way name resolution is typically handled for private end users is that they
send all their lookups to one DNS server that they were made aware of by their
internet service provider. When they communicate with another host for the first
time, e.g. to access a web site or send an email, they send a DNS request to
that server. The response is then likely cached for a while by their
operating system or their browser, but the DNS server still gets at least an
initial request and knows which host they are likely communicating with. Thus,
DNS requests are privacy-sensitive data that allow the server to build user
profiles.

Just encrypting traffic with the DNS server of course doesn't help, the problem
isn't eavesdroppers on the connection, it's all of a user's DNS meta data
heaping up at one central server. The operator of the DNS server doesn't even
have to have malicious intentions, or in fact any intentions to use the data.
Simply the fact that data from many users accumulates there makes the server a
worthwhile target for infiltration by malicious agents.

Preventing a buildup of large amounts of privacy-sensitive data on relatively
centralized servers is the motivation behind this thesis. The chosen approach is
to spread out name-lookup requests to many servers rather than using just one
resolver. This doesn't make incursions on users' privacy impossible, but should
significantly increase the work needed to gather a relevant amount of data. The
thinking is that a foreign party having access to a little bit of
privacy-sensitive data is relatively benign, and they are only able to build a
meaningful profile once they have acquired a considerable volume. With many
lookup servers available, peers are able to choose themselves how far they want
to spread out their requests.

\section{Goal}
There needs to be a great number of servers for this to work. This thesis
introduces a system in which these servers are the nodes of a \ac{DHT}, and the
mapping from name to network address the \emph{records} stored in it. To ensure
their great number, the main focus will be on incentivizing every user of the
lookup service to not just use it, but to become a \emph{peer} (i.e. a node of
the \ac{DHT}) that also contributes resources to providing it.

A good motivating use case, and one that's useful to keep in mind, is a
peer-to-peer chat system. Every peer has a fixed \emph{ID} by which he is known
to the others, but the network addresses can change frequently. The \ac{DHT} is
used to store the mapping from ID to network address. In fact, the rest of the
thesis assumes that each peer stores exactly one record containing his
\emph{reachability information} in the \ac{DHT}, where the key is equal to his
own peer ID (the key of a record is also referred to as the ID of the record).

Since network addresses change, peers frequently need to look up the current
addresses of others they want to chat with. Sending all these lookups to one
central server would give that server a pretty good picture of what a peer is
doing. Spreading them throughout the \ac{DHT} means a lot of people know a
little, but no single party knows enough to be useful.

Accessing a record is somewhat latency sensitive. While it is not a real-time
application, it becomes annoying if the lookup before sending a message takes
too long. Furthermore, it must be possible to update records, and for these
updates to be disseminated throughout the \ac{DHT} quickly. But we can assume
that the records are fairly small, containing just an ID, an address, and maybe
some additional information.

The system is likely capable of supporting other applications than such a chat
system, ones with fewer limiting assumptions. But they are out of scope of this
thesis.

Another underlying assumption is that the users are selfish. They want to expend
as little storage space, CPU time and network bandwidth as possible. But they
want to use the service and enjoy the benefit of having many peers to choose
from to do their lookups. Without any addition to the \ac{DHT}, they can do that
whether or not they contribute themselves; one more node is not going to make a
big difference in the grand scheme, and certainly not provide them a benefit.
This is the point of attack of the system: peers who aren't contributing must be
given a worse quality of service.

- TODO incorporate the following into goal section, along with other possible
  attacks?
The system should prevent free riding and be resistant to whitewashing.
Additionally, it should offer some protection against colluding peers trying to
appear as cooperating peers when they are not, and against Sybil attacks (TODO
ref background section for explanation).

\section{Necessity of Incentives}
The assumption that all peers are entirely selfish and need to be made to see a
benefit to themselves for every action required of them is very pessimistic. It
begs the question whether such a system is even necessary. Storing a few small
records containing reachability information and responding to queries concerning
them only takes little resources. Given that, it's not unreasonable to be
optimistic and believe that people wishing to use such a service are willing to
contribute to it.

In fact, \cite{mcgee2004gifting} surveys the motivation of users of a
peer-to-peer file sharing service to "gift", i.e. to seed without reward. The
authors find that there are users motivated by ideology ("information wants to
be free"), as well as ones acting altruistically, i.e. for whom the act of
sharing files itself is the motivation.

The later paper \cite{ripeanu2006gifting} examines motivations for gifting in
BitTorrent. The authors find that social characteristics strongly influence
sharing behavior, a user base that has a culture of sharing outside the file
sharing application will gift more in file sharing. They also find that a small
subset of the most active uploaders is responsible for a majority of the uploads
(in 95\% of torrents, over half the file is uploaded by the top 10\% of
uploaders).

This suggests that, while gifting, or generous behavior, in peer-to-peer
applications exists, relying on it may not be suitable for the system proposed
in this thesis. It could be suggested to peers that participating in offering
the service makes it better, in hopes of inviting generous peers' contributions,
as proposed in \cite{arma2009incentive_tor} for Tor. The problem is that this
may mostly just encourage the small subset of the most generous peers. But the
system absolutely needs many peers participating so users can spread out their
queries. Few peers doing most of the work is actually counterproductive to the
privacy goal.

One could also argue that it should be sufficient to write an implementation of
the \ac{DHT} in which peers contribute by default, perhaps with no option of
disabling it. The overhead of this is small, so users wouldn't change the
software for a small benefit. \cite{ripeanu2006gifting} even recognizes the
difficulty of configuring a BitTorrent client to freeride as a contribution to
the low levels of freeriding. But then a problem will arise if an alternative
implementation leaves out the contribution, e.g. out of laziness or for security
reasons (less code means less space for vulnerabilities).

While there is no certainty that incentives are absolutely necessary, there are
some indications that they are. \cite{cox2002serving}, which explores the
possibility of serving DNS via peer-to-peer also states "We need to find models
in which people have incentives to run servers rather than just take free rides
on others’ servers.". In the end, it may be best just to be prepared for the
worst.

\section{Outline}
- TODO check this for accuracy
- TODO is this necessary?

Chapter~\ref{chap:background} gives brief introductions into techniques and
concepts that are the basis of the work presented in this thesis.

Chapter~\ref{chap:system_description} describes the incentive system in detail.

Chapter~\ref{chap:implementation} describes the implementation of a simplified
implementation of the system, and a simulation used to evaluate it.

Chapter~\ref{chap:rep_avail} presents an evaluation of the system with regard to
the ability of cooperative peers to quickly establish themselves in the system
and enjoy use of the service without hindrance.

Chapter~\ref{chap:performance} examines the performance achievable by the system
theoretically and compares it with alternatives.

Chapter~\ref{chap:conclusion} concludes and gives an outlook towards interesting
open problems concerning the system.

\section{Related Work}
- TODO

The very initial motivation of this thesis was to address the privacy problem a
user has when he sends all his name resolution requests to a single server. This
problem goes away entirely if the server answers the requests satisfactorily but
is unable to tell which name the user queried for. This is the essence of
private information retrieval (PIR), \cite{gasarch2004survey} gives an overview.

Trivially, PIR is achieved if the server responds with the entire database, no
matter the query. More complex PIR schemes offer less prohibitively expensive
communication complexity, even more so computationally private information
retrieval (cPIR) like \cite{cachin1999computationally}, a relaxation using the
assumption attackers can only perform polynomial computations. However, even
they are impractical for the purpose of this thesis, since the server must
process the entire database for every query. If he didn't he could infer that
the record the user was looking for was in the part he didn't process.

DNS over TLS (DoT)\cite{RFC7858} encrypts traffic to DNS servers. As stated
previously, this doesn't solve the problem of privacy-sensitive data
accumulating in a central location.

DNS over HTTPS (DoH)\cite{RFC8484} does this as well, but has the potential to
also address the other problem. In conjunction with HTTP/2's server
push\cite{RFC7540}, a web server serving a web page can proactively include DNS
resolutions for the domain names linked in the page. The user then doesn't need
to resolve them himself. But this only really works for web browsing or similar
applications, where the server can reasonably anticipate which DNS resolutions a
client needs. It doesn't transfer to the chat system example from earlier.

The \ac{DHT} used in this thesis is a simplified variant of
P-Grid\cite{aberer2001pgrid}. One of the central promises of P-Grid is its
self-organizing load balancing, achieved by making peer identifiers independent
of the records a peer stores (the peer's path), as well as a cooperative
distributed algorithm extending or shrinking a peer's responsibility. This
functionality is not used in this thesis, but it is potentially useful in future
work.

The better known Kademlia\cite{maymounkov2002kademlia} is of a similar style, in
that routing takes place along a binary tree.

- various trust management
- various incentives
- freeriding, whitewashing, sybil attack etc.
- peers with utility functions that have a price for privacy
- that paper that found not all peers to be selfish
- that paper that looked at patterns of collusion
- that paper that found trusting peers initially, based on the trustworthiness
  of recent joiners, is sane

The NICE system presented in \cite{sherwood2006cooperative} is a distributed
trust inference system for peer-to-peer applications in general. It views them
as a set of two-party interactions, after each of which the two peers exchange a
signed statment, called cookie, about the quality of the interaction, basically
giving a rating. These can later be used by peers to prove to others that they
had a satisfactory interaction before and should therefore trust them now. If no
direct interaction took place, a chain of trust through mutual acquaintances can
also be presented as evidence of trustworthiness.

The purpose of NICE is to ensure that each peer only interacts with other
cooperative peers, whereas this thesis requires the system to also make sure
there are many cooperative peers to choose from. Otherwise, in an extreme case,
one peer could serve all queries and collect everyone's data.

- treat-before-trick
- gifting technologies paper (already mentioned above)
